---
title: "Activity Monitoring Data"
output: html_document
---


### table of content


- part 1: Mean total number of steps of taken per day
- part 2: Average daily activity pattern
- part 3: dataset with missing value
- part 4: difference weekdays and weekend




```{r Obtain data, echo=TRUE}
# Read data Activity dataset        
        path <- "C:/Users/vobis/Desktop/Coursera/5. Reproducible Resreach/project 1"
        name <- "activity.csv"
        dt_activity <- data.frame()
        dt_activity <- read.csv(paste(path, name, sep="/"), sep = ",")
```

#### part 1: Mean total number of steps of taken per day



```{r Part 1 Mean total step by day, echo=TRUE}
# Part 1 - Plot and mead/Medain off step by day      
        # Copy data.frame in new data.frame
                dt_activity_1 <- dt_activity        
        # Subset, exclude NA rows. 
                dt_activity_1 <- dt_activity_1[complete.cases(dt_activity_1),]
        # aggrate by day
                dt_steps_by_day <- as.data.frame(xtabs(steps ~ date  , dt_activity_1))      
        # plot Histogram        
                hist(dt_steps_by_day$Freq )
        # Calcualete mean and Median
                mean_activity_1 <- mean (dt_steps_by_day$Freq)
                median_activity_1 <- median(dt_steps_by_day$Freq)

        print(c("Mean:", round(mean_activity_1,2) ,"Median : " , round(median_activity_1,2)))
```

#### part 2: Average daily activity pattern


```{r Average daily Activity pattern, echo=TRUE}
# Part 2 - plot average 5 Min interval
        # Copy data.frame in new data.frame, keep raw date original 
                dt_activity_2 <- dt_activity        
        # Subset, exclude NA rows. 
                dt_activity_2 <- dt_activity_2[complete.cases(dt_activity_2),]
        # aggrate by day by mean
                dt_steps_by_day2 <- as.data.frame(aggregate(dt_activity_2$steps, by=list(dt_activity_2$interval), FUN=mean, data = dt_activity_2 ))
        # Max Value of mean
                maxvalue <- max(dt_steps_by_day2$x)
                dt_steps_by_day2_max <- dt_steps_by_day2[dt_steps_by_day2$x ==  maxvalue, 1]
        # plot
                plot_part2 <- plot(x=dt_steps_by_day2$Group.1 , y= dt_steps_by_day2$x , type = "l", main = "Average steps taken for each 5 min interval, Missing value are excluded")
                text_part2 <- paste("On average, in the next 5 min interval the most steps were taken;", round(dt_steps_by_day2_max,2), ", with ", round(maxvalue,2) ,sep = " ")

                print(text_part2)
                # additional comment. 
                print("Between8 and 9 in the morning, the most steps are taken")

```



#### part 3: dataset with missing value



```{r Dateset with missing value, echo=TRUE}
# part 3 - missing Value's

        # report number of missing NA, create new date to keep raw date original
        # and count number of row with missing data
                dt_activity_na <- dt_activity[is.na(dt_activity[,1]),]
                missing_na <- length(dt_activity_na[,1])
        # filing in missing Value's
                # value for missing values
                        mean_not_missing <- mean(dt_activity_1[complete.cases(dt_activity_1),1])
                # replece NA's with mean, create new dt for replaced data, so raw date will be original 
                        dt_replace_na <- dt_activity
                        dt_replace_na[is.na(dt_replace_na[,1]),1] <- mean_not_missing

        # Mean and Medain steps by each day
                        dt_steps_by_day_replaced_na <- as.data.frame(xtabs(steps ~ date  , dt_replace_na))      
        # Calcualete mean and Median, and max
                mean_activity_replaced_na <- mean (dt_steps_by_day_replaced_na$Freq)
                median_activity_replaced_na <- median(dt_steps_by_day_replaced_na$Freq)
                
        # plot average 5 min interval
                # aggrate by day, mean
                        dt_replace_na_agg <- as.data.frame(aggregate(dt_replace_na$steps, by=list(dt_replace_na$interval), FUN=mean, data = dt_replace_na))

                        max_activity_replaced_na <- max(dt_replace_na_agg$x)
                # text output - Part 1
                        text_part3_1_1 <- paste("Total number of NA, " , missing_na , sep = "")
                        text_part3_1_2 <- paste("Total missing rows, is" , round(missing_na/17568,2)*100 , "% of all observatoins" , sep = "")
                                
                                print(text_part3_1_1)
                                print(text_part3_1_2)
                # plot
                        plot(x=dt_replace_na_agg$Group.1 , y= dt_replace_na_agg$x , type = "l", main = "Average steps taken for each 5 min interval, Missing value are replaced")

                # text output - Part 2
                        text_part3_2 <- paste("NA replaced with mean steps(exclude NA)", round(mean_not_missing,2) ,sep = "")
                        text_part3_3 <- paste("Ater replace NA with mean, mean; ", round(mean_activity_replaced_na,2), " and median; ", round(median_activity_replaced_na,2) , "  number steps in highest interval; " , round(max_activity_replaced_na,2) , sep = "")     
                                print(text_part3_2)
                                print(text_part3_3)

                # text output - Part 3 [difference between exclude and replace]
                        Diff_mean <- round(mean_activity_replaced_na - mean_activity_1,0)
                        Diff_median <-round(median_activity_replaced_na - median_activity_1,0)
                        Diff_max <- round(max_activity_replaced_na - maxvalue,0)
                        text_part3_4 <- paste("Difference mean: ", Diff_mean , " | Difference median:" , Diff_median, " | Difference max steps in interval" , Diff_max ,  sep = "")
                        print(text_part3_4)
                        print("Due replacing Missing value, instead of exclude then, 13% of the total. The next things are changed;                              1. The patron on the plots look the same, only the y scale is less high. The highest number steps in an interval has drop with 22                              2. The mean and median has raised")
        
        ```

####  part 4: difference weekdays and weekend



```{r Difference weekdays and weekendays, echo=TRUE }
# part 4 - week /  weekend days

        # Add day name to data.frame
                dt_replace_na$day <- weekdays(as.Date(dt_replace_na$date), abbreviate = FALSE)
        # split dataset into weekday and weekenddays
                weekend  <- dt_replace_na[dt_replace_na$day %in% c("zaterdag","zondag"),]
                week     <- dt_replace_na[!(dt_replace_na$day %in% c("zaterdag","zondag")),]
        # aggrate by day, by mean
                dt_replace_na1_weekend <- as.data.frame(aggregate(weekend$steps, by=list(weekend$interval), FUN=mean, data = weekend ))
                dt_replace_na1_week <- as.data.frame(aggregate(week$steps, by=list(week$interval), FUN=mean, data = week ))
        # plot
                par(mfrow = c(2,1))
                plot(x=dt_replace_na1_weekend$Group.1 , y = dt_replace_na1_weekend$x , type = "l", main = "Weekend")
                plot(x=dt_replace_na1_week$Group.1 , y = dt_replace_na1_week$x , type = "l", main = "Week")

```


Difference between week and weekends.

1. During the week, their is one big pike between 8 and 9 in the morning. In the weekend is more spread over the day. 

